Efficiency

### 2018-01-08 DraftExec.Rnw
add imgs and rnw

\usepackage{nameref}

@@

\begin{figure}
\includegraphics[width=\textwidth]{XenonPoison.png}
\caption{Xenon poisoning effect following shutdown from \parencite{garland2005}}
\label{fig:XenonPoison.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{BoundaryLayer.png}
\caption{The development of a boundary layer as shown in \parencite{mollard2011}}
\label{fig:BoundaryLayer.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{LaminarTurbulent.png}
\caption{A simple comparison of laminar and turbulent flows from \parencite{NPdotnet2017}}
\label{fig:LaminarTurbulent.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{EarlyCavitation.JPG}
\caption{Cavitation bubbles form and collapse creating noise and some tiny bubbles long before boiling occurs}
\label{fig:EarlyCavitation.JPG}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{BoilingCavitation.JPG}
\caption{Only at very high temperatures do the bubbles endure at full-size in the water and reach the surface}
\label{fig:BoilingCavitation.JPG}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.7\textwidth]{IncreasingCavitation.png}
\caption{Cavitation frequently occurs first at the extremities, then spreads inwards across the blad face, as shown in \parencite{mollard2011}}
\label{fig:IncreasingCavitation.png}
\end{figure}

\begin{figure}
\includegraphics[width=0.7\textwidth]{CavitationTypes.png}
\caption{Different types of cavitation can occur on different parts of the propulsion system as shown in \parencite{carlton2007}}
\label{fig:CavitationTypes.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{SeparatedAerofoil.jpg}
\caption{Flow separation occurring over the top of an aerofoil in a wind tunnel. Image courtesy of Deutsches Zentrum fuer Luft- und Raumfahrt e. V. (DLR)}
\label{fig:SeparatedAerofoil.jpg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{FlowSeparation.png}
\caption{Flow separation involves flow reversal near a surface when a fluid is moving against an adverse pressure gradient, as shown in \parencite{mollard2011}}
\label{fig:FlowSeparation.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{PropellerAction.png}
\caption{Propellers generate thrust by accelerating a column of water, as shown in  \parencite[247]{mollard2011}}
\label{fig:PropellerAction.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{SkewRake.png}
\caption{The development of a boundary layer \parencite[262]{mollard2011}}
\label{fig:SkewRake.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{HelicalLine.png}
\caption{A blade section of a propeller traces a helical line around a cylinder as it is rotated \parencite{carlton2007}}
\label{fig:HelicalLine.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{PitchDefinition.png}
\caption{The definition of pitch is the distance traveled by a blade section along a cylinder, as given in \parencite{carlton2007}}
\label{fig:PitchDefinition.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{CPPropeller.png}
\caption{Controllable Pitch Propellers (CP Propellers) allow the pitch of the blades to be varied. Image credit: \parencite{mollard2011}}
\label{fig:CPPropeller.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{OpenWater.png}
\caption{A typical chart showing open water efficiency of a propeller, in this case a Wageningen B5-75. Image: MARIN}
\label{fig:OpenWater.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{SlipRatio.png}
\caption{The slip ratio of a propeller. Image credit: \cite{MAN2017}}
\label{fig:SlipRatio.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{AccelDecel.png}
\caption{he two types of ducted propeller, accelerating (left) and decelerating (right) as shown in \parencite[249]{mollard2011}}
\label{fig:AccelDecel.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{DuctStreamlines.png}
\caption{The streamlines of flow through different duct types \parencite[214]{lewis1988}}
\label{fig:DuctStreamlines.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{CavitatingDuct.png}
\caption{An example of the cavitation occurring at the blade tips of a ducted propeller. Image credit \cite{moulijn2015}}
\label{fig:CavitatingDuct.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{DuctedEfficiency.png}
\caption{Accelerating ducts produce a net positive thrust at low advance ratios, but a net negative thrust at higher advance ratios as the drag on the nozzle exterior increases. Image credit:Open-water test results of Ka 3-65 screw series with nozzle no. 19A \parencite{oosterveld1970}}
\label{fig:DuctedEfficiency.png}
\end{figure}

@@

In order to achieve a reduction in cavitation, however, care must be taken in the design of a decelerating duct.  Since the negative thrust produced by the duct must necessitate additional work being done by the impeller, there can be circumstances where the additional loading of the impeller out-weighs the impacts of elevating the pressure around it. \cite[24-25]{oosterveld1970} provides a more thorough derivation of the circumstances in which this can be reliably achieved, which generally include larger blade area ratios or with more rows of rotors for higher loading, or circumstances where the loading of the propeller is quite low.  This creates an inevitable and necessary tension between optimisation of a duct for propulsive efficiency and reduction of cavitation, which will be returned to in Section \ref{waterjets} (\nameref{waterjets}).

@@

\begin{figure}
\includegraphics[width=\textwidth]{PumpjetOutline.png}
\caption{In \cite{carlton2007} the pump jet is described as a development of a classical ducted propeller, with the noticeable addition of a stator}
\label{fig:PumpjetOutline.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{PumpjetWaterjet.png}
\caption{A comparison of a pumpjet and waterjet as given in \cite{wislicenus1973}}
\label{fig:PumpjetWaterjet.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{EfficiencyJapan.png}
\caption{A graph from a Japanese study of waterjet efficiency shows efficiency rapidly declining below craft velocities of 10m/s, in accordance with theory. \parencite{fujisawa1995}}
\label{fig:EfficiencyJapan.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{EfficiencyTransport.png}
\caption{An Australian experimental study aimed at determining the crossover point for the efficiency of jets and propellers in the 20-30kt range revealed much higher power was required across the entire speed range \parencite{kamal2015}}
\label{fig:EfficiencyTransport.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{BMTShape.png}
\caption{The 'fully submerged waterjet' proposed by BMT and Rolls Royce for military surface craft may more closely represent the duct shape used on submarines than waterjets that release above the surface. \parencite{giles2010}}
\label{fig:BMTShape.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{BMTEfficiency.png}
\caption{Studies on submerged waterjets for military surface craft currently show that the efficiency of jets decline sharply below 10kt \parencite{giles2010}}
\label{fig:BMTEfficiency.png}
\end{figure}

@@

There are other some fundamental limitations on the peak efficiency that can be achieved by jets based on simple calculations of ideal efficiency from momentum theory, which are discussed in Section \ref{tradeoffs} (\nameref{tradeoffs}).  They also provide enduring, reliable limitations on the extent to which ingenuity or design changes might be able to improve the efficiency of jets at low speed. But the primary reason why jet efficiencies decline towards zero at low speed is to do with flow separation \parencite{fujisawa1995}, which is what occurs when the assumptions inherent in momentum theory's analysis of an 'ideal' actuator disc start to break down.

@@

As can be seen in figure \ref{fig:JetHighFlow.png}, the duct of a jet is designed to operate by deflecting the flow away from the natural path that would occur for a propeller, which would accelerate the stream over its blades directly, with the flow necessarily contracting with the acceleration as shown in Figure \ref{fig:PropellerAction.png}.   Instead, pumpjets tend to use some moderate expansion of the flow in advance of the impeller (diffusion) in order to raise its pressure, and reduce its velocity.  (Allowing the blades to do work on the water at lower speeds and higher pressure is the essence of what reduces cavitation.)  The blades then do their work on the liquid, further raising its pressure.  The duct allows this pressure then to be converted back to kinetic energy, as the water accelerates to move through the contracting area in the nozzle.  This difference in the shape of the flow, and the consequent impacts on the static pressures, are demonstrated in Figure \ref{fig:FlowComparison.png} taken from \cite{mccormick1963design}.

\begin{figure}
\includegraphics[width=\textwidth]{FlowComparison.png}
\caption{The effect of the duct is to deflect the flow away from the streamlines of an open propeller, elevating the static pressure at the stator blades by slowing the flow down.}
\label{fig:FlowComparison.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{JetHighFlow.png}
\caption{At high flow rates the water has sufficient velocity (indicated by arrow length) to overcome the adverse pressure gradients created by the duct shape, shown by the change from green to red. After the impeller has added further pressure to the water, the water accelerates out of the contracting nozzle through the favourable pressure gradient (red to green) to produce a jet, reaching equal pressure to outside at the exit}
\label{fig:JetHighFlow.png}
\end{figure}

@@

Flow separation also tends to lead to unsteady flows, with vortices periodically being shed into the main stream as they are speed up or change shape, and then ingested by the impeller and stator blades.  As can be seen by the diagram in Figure \ref{fig:JetLowFlow.png}, regions abound which are adjacent to or behind flow blockages or reversing flow, where considerable turbulence and unsteady forces would exist.  These effects also have acoustic consequences, which will be discussed in Section \ref{acousticconsiderations} (\nameref{acousticconsiderations}).

\begin{figure}
\includegraphics[width=\textwidth]{JetLowFlow.png}
\caption{At low flow rates the water has insufficient velocity to overcome the adverse pressure gradients created by the duct shape, and flow separation begins to occur on the duct intake, and between the blades beginning near the roots (stall).  Flow through the jet would be unsteady, with excessive energy expended creating vortices and turbulence that results in little thrust.}
\label{fig:JetLowFlow.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{Comparison.png}
\caption{At the end of a tapering body the impact of the contracting hub will be much larger near the intake than near the tail, making the degree of diffusion ahead of the impeller, as well as contraction of the flow in the nozzle after it, much larger in impact than it appears by looking at the external shape. Credit \cite{gearhart1966selection}}
\label{fig:Comparison.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{Triomph.jpg}
\caption{The outer shape of pumpjets on the aft end of torpedoes or submarines, such as this one on the French Triomphant class on which the Barracuda's jet will be based, may understate the degree of deceleration that the duct achieves. Credit \cite{reddit2016}}
\label{fig:Triomph.jpg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{PumpjetEfficiency.png}
\caption{Ideal jet efficiencies sharply appoach zero at low values of $\Delta V / V_0$, where as ideal propeller efficiencies without a duct $K = 0$ approach unity. Credit: \cite{wislicenus1973}}
\label{fig:PumpjetEfficiency.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{EfficiencyBrandau.png}
\caption{Alternative derivations of ideal jet efficiency also show sharply declining efficiency for low values of $\Delta V / V_0$ Credit: \cite{brandau1967}}
\label{fig:EfficiencyBrandau.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{EfficiencyMollard.png}
\caption{Other modern reference texts might demonstrate jet or propeller efficiencies with different x-axes, effectively reversing the way the graph should be understood for a single system operating at different speeds. Credit:  \cite{mollard2011}}
\label{fig:EfficiencyMollard.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{EfficiencyBulten.png}
\caption{Another graph representing efficiency as a function of the ratio of jet velocity to incoming velocity, which shows efficiency declining at low thrust (to the right).  The impact of increasing duct losses can be clearly seen, as opposed to the 'ideal' propeller without a duct. Credit: \cite{bulten2006numerical}}
\label{fig:EfficiencyBulten.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{PropDuct.png}
\caption{Even the addition of an accelerating ducts substantially reduces efficiency of propellers at very low loadings, or for constant loads, low speeds.  Maximum attainable efficiency is also reduced. Their benefit lies in the bulging of the curve at low advance ratios (higher loads).  Credit: \cite[6]{willemsen2013}}
\label{fig:PropDuct.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{PumpUp.png}
\caption{The mysterious efficiency comparison chart which has been doing the rounds in the Australian debate probably quite reasonably reflects theoretical physical limits of 'ideal' pumpjets and propellers. Credit: \cite{stanford2017}}
\label{fig:PumpUp.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{stall.png}
\caption{Flow separation, or stall, can occur when the flow over the blades of a pump slows too much, as shown in \cite{li2013}}
\label{fig:stall.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{backflow.png}
\caption{Stall results in water flowing backwards near the hub of the blades at low flow rates, as shown in \cite{li2014rotating}}
\label{fig:backflow.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{AcceleratingIdeal.png}
\caption{Calculations of 'ideal' efficiencies of accelerating ducts (neglecting drag) in \cite{oosterveld1970} show that a duct can only improve the efficiency of a propeller if it is loaded above a certain degree, and that longer ducts produce better improvements in 'ideal' efficiencies}
\label{fig:AcceleratingIdeal.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{FineOptimised.png}
\caption{Extensive work has been done exploring compromise designs for decelerating ducted propellers (pumpjets), including with very modest deceleration in the intake, limited shroud taper, short duct lengths, or essentially incporate some of the advantages of accelerating ducts. Credit: \cite{abdel2010}}
\label{fig:FineOptimised.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{FineEfficiency.png}
\caption{In experimental testing such designs still exhibit sharply falling effiency at high advance ratios, or low speed. Credit:  \cite{abdel2010}}
\label{fig:FineEfficiency.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{FineCavitation.png}
\caption{TAt optimally efficient conditions (corresponding to 27kt) such designs still exhibit some cavitation. Credit: \cite{abdel2010}}
\label{fig:FineCavitation.png}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{LowIVR.png}
\caption{A an intake may experience risks of flow separation even at low speeds (Inlet Velocity Ratio) on some surfaces due to certain flow deflections being more severe. Source: \cite{bulten2006numerical}}
\label{fig:LowIVR.png}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{HighIVR.png}
\caption{A an intake may experience risk of flow separation on different surfaces at high speeds (Inlet Velocity Ratio) where other flow deflections are more severe. Source: \cite{bulten2006numerical}}
\label{fig:HighIVR.png}
\end{figure}

>>>>

%  For caption width
\usepackage{graphicx,caption}
% For Centering figs
\usepackage[export]{adjustbox}
% For split figures
\usepackage{subcaption}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{XenonPoison.png}
\caption{Xenon poisoning effect following shutdown from \parencite{garland2005}}
\label{fig:XenonPoison.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{BoundaryLayer.png}
\caption{The development of a boundary layer as shown in \parencite{mollard2011}}
\label{fig:BoundaryLayer.png}
\end{figure}

@@

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.7\linewidth, center]{Laminar.png}
\caption{Laminar}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.7\linewidth, center]{Turbulent.png}
\caption{Turbulent}
\end{subfigure}

\caption{A simple comparison of laminar and turbulent flows from \parencite{NPdotnet2017}}
\label{fig:LaminarTurbulent.png}
\end{figure}

@@

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth, height=5cm]{EarlyCavitation.JPG}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth, height=5cm]{BoilingCavitation.JPG}
\caption{}
\end{subfigure}

\caption{(a) Cavitation bubbles form and collapse creating noise and some tiny bubbles long before boiling occurs. (b) Only at very high temperatures do the bubbles endure at full-size in the water and reach the surface.}
\label{fig:EarlyCavitation.JPG}
\end{figure}

@@

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{IncreasingCavitation.png}
\caption{Cavitation frequently occurs first at the extremities, then spreads inwards across the blad face, as shown in \parencite{mollard2011}}
\label{fig:IncreasingCavitation.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{CavitationTypes.png}
\caption{Different types of cavitation can occur on different parts of the propulsion system as shown in \parencite{carlton2007}}
\label{fig:CavitationTypes.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{SeparatedAerofoil.jpg}
\caption{Flow separation occurring over the top of an aerofoil in a wind tunnel. Image courtesy of Deutsches Zentrum fuer Luft- und Raumfahrt e. V. (DLR)}
\label{fig:SeparatedAerofoil.jpg}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{FlowSeparation.png}
\caption{Flow separation involves flow reversal near a surface when a fluid is moving against an adverse pressure gradient, as shown in \parencite{mollard2011}}
\label{fig:FlowSeparation.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{PropellerAction.png}
\caption{Propellers generate thrust by accelerating a column of water, as shown in  \parencite[247]{mollard2011}}
\label{fig:PropellerAction.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{SkewRake.png}
\caption{The development of a boundary layer \parencite[262]{mollard2011}}
\label{fig:SkewRake.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{HelicalLine.png}
\caption{A blade section of a propeller traces a helical line around a cylinder as it is rotated \parencite{carlton2007}}
\label{fig:HelicalLine.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{PitchDefinition.png}
\caption{The definition of pitch is the distance traveled by a blade section along a cylinder, as given in \parencite{carlton2007}}
\label{fig:PitchDefinition.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.8\textwidth}
\includegraphics[width=0.8\textwidth, center]{CPPropeller.png}
\caption{Controllable Pitch Propellers (CP Propellers) allow the pitch of the blades to be varied. Image credit: \parencite{mollard2011}}
\label{fig:CPPropeller.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.8\textwidth}
\includegraphics[width=0.8\textwidth]{OpenWater.png}
\caption{A typical chart showing open water efficiency of a propeller, in this case a Wageningen B5-75. Image: MARIN}
\label{fig:OpenWater.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.8\textwidth}
\includegraphics[width=0.8\textwidth, center]{SlipRatio.png}
\caption{The slip ratio of a propeller. Image credit: \cite{MAN2017}}
\label{fig:SlipRatio.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=\textwidth}
\includegraphics[width=\textwidth]{AccelDecel.png}
\caption{The two types of ducted propeller, accelerating (left) and decelerating (right) as shown in \parencite[249]{mollard2011}}
\label{fig:AccelDecel.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{DuctStreamlines.png}
\caption{The streamlines of flow through different duct types \parencite[214]{lewis1988}}
\label{fig:DuctStreamlines.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{CavitatingDuct.png}
\caption{An example of the cavitation occurring at the blade tips of a ducted propeller. Image credit \cite{moulijn2015}}
\label{fig:CavitatingDuct.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{DuctedEfficiency.png}
\caption{Accelerating ducts produce a net positive thrust at low advance ratios, but a net negative thrust at higher advance ratios as the drag on the nozzle exterior increases. Image credit:Open-water test results of Ka 3-65 screw series with nozzle no. 19A \parencite{oosterveld1970}}
\label{fig:DuctedEfficiency.png}
\end{figure}

@@

In order to achieve a reduction in cavitation, however, care must be taken in the design of a decelerating duct.  Since the negative thrust produced by the duct must necessitate additional work being done by the impeller, there can be circumstances where the additional loading of the impeller out-weighs the impacts of elevating the pressure around it. \cite[24-25]{oosterveld1970} provides a more thorough derivation of the circumstances in which this can be reliably achieved, which generally include larger blade area ratios or with more rows of rotors for higher loading, or circumstances where the loading of the propeller is quite low.  This creates an inevitable and necessary tension between optimisation of a duct for propulsive efficiency and reduction of cavitation, which will be returned to in Section \ref{waterjets}.

@@

\begin{figure}[h]
\captionsetup{width=0.4\textwidth}
\includegraphics[width=0.4\textwidth, center]{PumpjetOutline.png}
\caption{In \cite{carlton2007} the pump jet is described as a development of a classical ducted propeller, with the noticeable addition of a stator}
\label{fig:PumpjetOutline.png}
\end{figure}

@@

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{Pumpjet.png}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{Waterjet.png}
\caption{}
\end{subfigure}

\caption{A comparison of a pumpjet (a) and waterjet (b) as given in \cite{wislicenus1973}}
\label{fig:PumpjetWaterjet.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{EfficiencyJapan.png}
\caption{A graph from a Japanese study of waterjet efficiency shows efficiency rapidly declining below craft velocities of 10m/s, in accordance with theory. \parencite{fujisawa1995}}
\label{fig:EfficiencyJapan.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.6\textwidth}
\includegraphics[width=0.6\textwidth, center]{EfficiencyTransport.png}
\caption{An Australian experimental study aimed at determining the crossover point for the efficiency of jets and propellers in the 20-30kt range revealed much higher power was required across the entire speed range \parencite{kamal2015}}
\label{fig:EfficiencyTransport.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=\textwidth}
\includegraphics[width=\textwidth]{BMTShape.png}
\caption{The 'fully submerged waterjet' proposed by BMT and Rolls Royce for military surface craft may more closely represent the duct shape used on submarines than waterjets that release above the surface. \parencite{giles2010}}
\label{fig:BMTShape.png}
\end{figure}

@@

\begin{figure}[p]

\begin{subfigure}[b]{0.9\textwidth}
\includegraphics[width=0.8\linewidth, center]{BMTEfficiency1.png}
\end{subfigure}
\begin{subfigure}[b]{0.9\textwidth}
\includegraphics[width=0.8\linewidth, center]{BMTEfficiency2.png}
\end{subfigure}

\caption{Studies on submerged waterjets for military surface craft currently show that the efficiency of jets decline sharply below 10kt \parencite{giles2010}}
\label{fig:BMTEfficiency.png}
\end{figure}

@@

There are other some fundamental limitations on the peak efficiency that can be achieved by jets based on simple calculations of ideal efficiency from momentum theory, which are discussed in Section \ref{tradeoffs}.  They also provide enduring, reliable limitations on the extent to which ingenuity or design changes might be able to improve the efficiency of jets at low speed. But the primary reason why jet efficiencies decline towards zero at low speed is to do with flow separation \parencite{fujisawa1995}, which is what occurs when the assumptions inherent in momentum theory's analysis of an 'ideal' actuator disc start to break down.

@@

As can be seen in Figure \ref{fig:JetHighFlow.png}, the duct of a jet is designed to operate by deflecting the flow away from the natural path that would occur for a propeller, which would accelerate the stream over its blades directly, with the flow necessarily contracting with the acceleration as shown in Figure \ref{fig:PropellerAction.png}.   Instead, pumpjets tend to use some moderate expansion of the flow in advance of the impeller (diffusion) in order to raise its pressure, and reduce its velocity.  (Allowing the blades to do work on the water at lower speeds and higher pressure is the essence of what reduces cavitation.)  The blades then do their work on the liquid, further raising its pressure.  The duct allows this pressure then to be converted back to kinetic energy, as the water accelerates to move through the contracting area in the nozzle.  This difference in the shape of the flow, and the consequent impacts on the static pressures, are demonstrated in Figure \ref{fig:FlowComparison.png} taken from \cite{mccormick1963design}.

\begin{figure}
\captionsetup{width=0.4\textwidth}
\includegraphics[width=0.4\textwidth, center]{FlowComparison.png}
\caption{The effect of the duct is to deflect the flow away from the streamlines of an open propeller, elevating the static pressure at the stator blades by slowing the flow down.}
\label{fig:FlowComparison.png}
\end{figure}

@@

\begin{figure}[p]
\captionsetup{width=1\textwidth}
\includegraphics[width=1\textwidth]{JetHighFlow.png}
\caption{At high flow rates the water has sufficient velocity (indicated by arrow length) to overcome the adverse pressure gradients created by the duct shape, shown by the change from green to red. After the impeller has added further pressure to the water, the water accelerates out of the contracting nozzle through the favourable pressure gradient (red to green) to produce a jet, reaching equal pressure to outside at the exit}
\label{fig:JetHighFlow.png}
\end{figure}

@@

Flow separation also tends to lead to unsteady flows, with vortices periodically being shed into the main stream as they are speed up or change shape, and then ingested by the impeller and stator blades.  As can be seen by the diagram in Figure \ref{fig:JetLowFlow.png}, regions abound which are adjacent to or behind flow blockages or reversing flow, where considerable turbulence and unsteady forces would exist.  These effects also have acoustic consequences, which will be discussed in Section \ref{acousticconsiderations}.

\begin{figure}[p]
\captionsetup{width=1\textwidth}
\includegraphics[width=1\textwidth]{JetLowFlow.png}
\caption{At low flow rates the water has insufficient velocity to overcome the adverse pressure gradients created by the duct shape, and flow separation begins to occur on the duct intake, and between the blades beginning near the roots (stall).  Flow through the jet would be unsteady, with excessive energy expended creating vortices and turbulence that results in little thrust.}
\label{fig:JetLowFlow.png}
\end{figure}

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{Comparison.png}
\caption{At the end of a tapering body the impact of the contracting hub will be much larger near the intake than near the tail, making the degree of diffusion ahead of the impeller, as well as contraction of the flow in the nozzle after it, much larger in impact than it appears by looking at the external shape. Credit \cite{gearhart1966selection}}
\label{fig:Comparison.png}
\end{figure}

\begin{figure}[p]
\captionsetup{width=0.8\textwidth}
\includegraphics[width=0.8\textwidth, center]{Triomph.jpg}
\caption{The outer shape of pumpjets on the aft end of torpedoes or submarines, such as this one on the French Triomphant class on which the Barracuda's jet will be based, may understate the degree of deceleration that the duct achieves. Credit \cite{reddit2016}}
\label{fig:Triomph.jpg}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.6\textwidth}
\includegraphics[width=0.7\textwidth, center]{PumpjetEfficiency.png}
\caption{Ideal jet efficiencies sharply appoach zero at low values of $\Delta V / V_0$, where as ideal propeller efficiencies without a duct $K = 0$ approach unity. Credit: \cite{wislicenus1973}}
\label{fig:PumpjetEfficiency.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.5\textwidth, center]{EfficiencyBrandau.png}
\caption{Alternative derivations of ideal jet efficiency also show sharply declining efficiency for low values of $\Delta V / V_0$ Credit: \cite{brandau1967}}
\label{fig:EfficiencyBrandau.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.6\textwidth}
\includegraphics[width=0.6\textwidth, center]{EfficiencyMollard.png}
\caption{Other modern reference texts might demonstrate jet or propeller efficiencies with different x-axes, effectively reversing the way the graph should be understood for a single system operating at different speeds. Credit:  \cite{mollard2011}}
\label{fig:EfficiencyMollard.png}
\end{figure}

\begin{figure}[h] \captionsetup{width=0.6\textwidth}
\includegraphics[width=0.8\textwidth, center]{EfficiencyBulten.png}
\caption{Another graph representing efficiency as a function of the ratio of jet velocity to incoming velocity, which shows efficiency declining at low thrust (to the right).  The impact of increasing duct losses can be clearly seen, as opposed to the 'ideal' propeller without a duct. Credit: \cite{bulten2006numerical}}
\label{fig:EfficiencyBulten.png}
\end{figure}

@@

\begin{figure}[h] \captionsetup{width=0.5\textwidth}
\includegraphics[width=\textwidth]{PropDuct.png}
\caption{Even the addition of an accelerating ducts substantially reduces efficiency of propellers at very low loadings, or for constant loads, low speeds.  Maximum attainable efficiency is also reduced. Their benefit lies in the bulging of the curve at low advance ratios (higher loads).  Credit: \cite[6]{willemsen2013}}
\label{fig:PropDuct.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.6\textwidth, center]{PumpUp.png}
\caption{The mysterious efficiency comparison chart which has been doing the rounds in the Australian debate probably quite reasonably reflects theoretical physical limits of 'ideal' pumpjets and propellers. Credit: \cite{stanford2017}}
\label{fig:PumpUp.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.6\textwidth}
\includegraphics[width=0.6\textwidth, center]{stall.png}
\caption{Flow separation, or stall, can occur when the flow over the blades of a pump slows too much, as shown in \cite{li2013}}
\label{fig:stall.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.8\textwidth}
\includegraphics[width=0.8\textwidth, center]{backflow.png}
\caption{Stall results in water flowing backwards near the hub of the blades at low flow rates, as shown in \cite{li2014rotating}}
\label{fig:backflow.png}
\end{figure}

@@

\begin{figure}[h] \captionsetup{width=0.5\textwidth}
\includegraphics[width=\textwidth]{AcceleratingIdeal.png}
\caption{Calculations of 'ideal' efficiencies of accelerating ducts (neglecting drag) in \cite{oosterveld1970} show that a duct can only improve the efficiency of a propeller if it is loaded above a certain degree, and that longer ducts produce better improvements in 'ideal' efficiencies}
\label{fig:AcceleratingIdeal.png}
\end{figure}

@@

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{FineOptimised.png}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{FineCavitation.png}
\caption{}
\end{subfigure}

\caption{(a) Extensive work has been done exploring compromise designs for decelerating ducted propellers (pumpjets), including with very modest deceleration in the intake, limited shroud taper, short duct lengths, or essentially incporate some of the advantages of accelerating ducts. (b) TAt optimally efficient conditions (corresponding to 27kt) such designs still exhibit some cavitation. Credit: \cite{abdel2010}}
\label{fig:FineOptimised.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{FineEfficiency.png}
\caption{In experimental testing such designs still exhibit sharply falling effiency at high advance ratios, or low speed. Credit:  \cite{abdel2010}}
\label{fig:FineEfficiency.png}
\end{figure}

@@

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{LowIVR.png}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{HighIVR.png}
\caption{}
\end{subfigure}

\caption{(a) A an intake may experience risks of flow separation even at low speeds (Inlet Velocity Ratio) on some surfaces due to certain flow deflections being more severe. (b) A an intake may experience risk of flow separation on different surfaces at high speeds (Inlet Velocity Ratio) where other flow deflections are more severe. Source: \cite{bulten2006numerical}}
\label{fig:LowIVR.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{DuctSeparation.png}
\caption{Flow separation occurs at on the outside of an accelerating duct at high $J$ values, which corresponds to low loads, or speeds. For high $J$ values, the pressure reduction on the inside of the duct, creating positive duct thrust, can be seen. Source: \cite[73]{willemsen2013}}
\label{fig:DuctSeparation.png}
\end{figure}
